---
title: "Strawberries: exploratory data analysis"
author: Jiaxin Li(U77902942)
date: 2023 Oct 23
format: pdf
engine: knitr
editor: 
  markdown: 
    wrap: 72
---

<!-- # Assignment -->

<!-- Using our class discussions and this document as a starting point, -->
<!-- produce an EDA report. The report should describe the data itself so -->
<!-- that readers understand the data sources used in the report and how you -->
<!-- cleaned and organized the data for analysis. -->

<!-- The sections below suggest how the report might be organized. The report -->
<!-- should be succinct, communicating the information that you believe will -->
<!-- be helpful to someone doing a fuller analysis of the data or using the -->
<!-- data for model building. Implementation details should be included in -->
<!-- commentary that is included in code. -->

<!-- Sections of the document as it was originally presented in class have -->
<!-- been commented so that you can see them in the code. -->

<!-- ## Data acquisition and assessment -->

<!-- -   Data sources\ -->
<!-- -   Assumptions and motivations -->

<!-- ## Data cleaning and organization -->

<!-- Outline the approach taked to clean and organize the data. -->

<!-- ## References -->

<!-- ### Material about strawberries -->

<!-- [WHO says strawberries may not be so safe for -->
<!-- you--2017March16](https://med.news.am/eng/news/13621/who-says-strawberries-may-not-be-so-safe-for-you.html) -->

<!-- [Pesticides + poison gases = cheap, year-round strawberries -->
<!-- 2019March20](https://www.ewg.org/foodnews/strawberries.php) -->

<!-- [Multistate Outbreak of Hepatitis A Virus Infections Linked to Fresh -->
<!-- Organic -->
<!-- Strawberries-2022March5](https://www.cdc.gov/hepatitis/outbreaks/2022/hav-contaminated-food/index.htm) -->

<!-- [Strawberry makes list of cancer-fighting -->
<!-- foods-2023May31](https://issuu.com/mechlocal/docs/053123_mech_asf/s/25386339) -->

<!-- ### Technical references -->

<!-- In their handbook ["An introduction to data cleaning with R" by Edwin de -->
<!-- Jonge and Mark van der -->
<!-- Loo](https://cran.r-project.org/doc/contrib/de_Jonge+van_der_Loo-Introduction_to_data_cleaning_with_R.pdf), -->
<!-- de Jonge and van der Loo go into detail about specific data cleaning -->
<!-- isssues and how to handle them in R. -->

<!-- ["Problems, Methods, and Challenges in Comprehensive Data Cleansing" by -->
<!-- Heiko Müller and Johann-Christoph -->
<!-- Freytag](https://www.researchgate.net/profile/Heiko-Mueller/publication/228929938_Problems_methods_and_challenges_in_comprehensive_data_cleansing/links/09e415101b58541e2c000000/Problems-methods-and-challenges-in-comprehensive-data-cleansing.pdf) -->
<!-- is a good companion to the de Jonge and van der Loo handbook, offering -->
<!-- additional insights. -->

<!-- ## Initial questions -->

<!-- -   Initial questions about strawberries, the data, and about the work -->
<!--     you are undertaking. -->

<!-- The questions: (1) What is the degree of harmfulness of each chemical, -->
<!-- which chemicals are more harmful and which are relatively harmless? (2) -->
<!-- How are strawberries sold in different states and in different years? -->
<!-- What about the use of chemicals? (3) Whether there is a correlation -->
<!-- between different variables, e.g., whether strawberry sales are related -->
<!-- to the year or the state in which they are located, etc. -->

<!-- ## The data -->

<!-- Describe the source and original condition of the data: organization, -->
<!-- problems with the data that needed to be addressed and so on. Cite data -->
<!-- sources. -->

<!-- The data set for this assignment has been selected from: -->
<!-- [USDA_NASS](https://quickstats.nass.usda.gov) <br> The data have been -->
<!-- stored on NASS here: -->
<!-- [USDA_NASS_strawb_2023SEP19](https://quickstats.nass.usda.gov/results/45FBC825-B104-38E2-9802-839F5F3C7036) -->

<!-- Make relevant observations in the document and in your code about data. -->
<!-- Add commentary to the code so that anthoer analysts could use or extend -->
<!-- your code. -->

<!-- Discusss missing data, inclding how you handled it. Be careful to point -->
<!-- out where NA's are being produced during processing and are not data -->
<!-- missing in the original data. -->

<!-- Where it is relevant, include information of how you have organized the -->
<!-- data for analysis. It might, for example, be helpful to know that there -->
<!-- is both agricultural census data and survey data. It might be helpful to -->
<!-- discuss data that appears to be redundant between these two sources. -->

<!-- Make sure you include details in your discussion and in your code about -->
<!-- other data and information you used in your work. Cite sources and -->
<!-- provide detail that would allow another analyst to reproduce your work. -->

```{r, include=FALSE}
#| label: Load libraries
#| warning: false
#| message: false
#| echo: false

library(knitr)  
library(kableExtra)
library(tidyverse)
library(stringr)
```

<!-- Read the file -->

```{r warning=FALSE, message=FALSE, include=FALSE}
#| label: read data - glimpse 
#| warning: false
#| message: false
#| echo: false

strawberry <- read_csv("strawberry(2).csv", col_names = TRUE)

# glimpse(strawberry)

```

```{r, include=FALSE}
#| label: drop one-item columns
#| echo: false

## define function
drop_one_value_col <- function(df){
col_name <- NULL
col_val <- NULL
suppressWarnings({
for(i in 1:dim(df)[2]){
if((df |> distinct(df[,i]) |> count()) == 1){
  col_name = c(col_name, colnames(df[i]))
  col_val = c(col_val, df[1,i])  
} }
})

if(is.null(col_name)){return("No Columns to drop")}else{
   col_val = unlist(col_val)
   attributes(col_val) = NULL
   drp = data.frame(col_name, col_val)
   return(drp)
   }
}

str <- drop_one_value_col(strawberry)

# str |> kable(caption = "Dropped Single-Value Columns: names and values")

str <- str$col_name

strawberry <- strawberry |> select(!all_of(str))



## applying the function a second time 
## tests the function when there aren't any 
## one-value columns
#####  drop_one_value_col(strawberry)
```

<!-- Glimpse of strawberry data after dropping single-value columns. -->

```{r, include=FALSE}
#| label: glimpse of strawberry data
#| echo: false

## glimpse(strawberry)

```

<!-- ## Examine the data. How is it organized? -->

<!-- ### Is every line associated with a state? -->

```{r, include=FALSE}
#| label: examine rows
#| echo: false

## is every line associated with a state?

## state_all contains the number of rows containing data 
## for each of the 47 strawberry-growing states.
state_all <- strawberry |> group_by(State) |> count()

## test if every row is associated with a state by summing the 
## counts and testing for equality with the total rows in the 
## data frame

# if(sum(state_all$n) == dim(strawberry)[1]){print("Every row has value in the State column.")}

```

<!-- ### Which state has the most rows? -->

```{r, include=FALSE}
#| label: which state has the most rows
#| echo: false

state_max <- state_all$State[which(state_all$n ==  max(state_all$n)  )]

```

<!-- The data is organized by state.   -->

<!-- The state with the most rows is `r paste(state_max)`. -->

<!-- ## Examine California data -->

```{r, include=FALSE}
#| label: examine California data
#| echo: false

## filter rows of California data from the CENSUS data
calif_census <- strawberry |> filter((State=="CALIFORNIA") & (Program=="CENSUS"))


## ## filter rows of California data from the SURVEY data
calif_survey <- strawberry |> filter((State=="CALIFORNIA") & (Program=="SURVEY"))

census_col <- colnames(calif_census)

survey_col <- colnames(calif_survey)

```

<!-- ### List the composite columns  -->

<!-- Census: `r paste(census_col[c(6, 8)])` -->

<!-- Survey: `r paste(survey_col[c(6,7,8)])` -->

<!-- ## Separate CENSUS and SURVEY into two Data Frames -->

<!-- In the strawberry data frame,  -->

<!-- The CENSUS rows contains marketing, sales, and productiong data.  The SURVEY rows contain rows which may be redundant with the CENSUS rows and chemical application rows. -->

<!-- After splitting CENSUS and SURVEY rows into two data frames,  -->

<!-- finish organizing the CENSUS data first.  Then, organize the -->

<!-- SURVEY data frame splitting the marketing, and production data from the chemical application data. -->

```{r, include=FALSE}
#| label: split srawberry into census and survey pieces
#| echo: false

strwb_census <- strawberry |> filter(Program == "CENSUS")

strwb_survey <- strawberry |> filter(Program == "SURVEY")

## check that all of the rows are accounted for

## nrow(strawberry) == (nrow(strwb_census) + nrow(strwb_survey))

## Move marketing-related rows in strw_b_chem 
## to strw_b_sales

## clean up the environment

rm(calif_census, calif_survey, state_all)

```

<!-- # Complete with the census data frame -->

<!-- ## Separate composite columns and clean the Value column -->

<!-- Composite columns in the strwb_census: Data Item, Domain category -->

<!-- Column separators in CENSUS: ",", "-", ":" -->

<!-- ### Separate `Data Item` into columns by "," -->

```{r, include=FALSE}
#| label: split Data Item
#| echo: false

## This will be done in stages --

####################################################
## split `Data Item` into "Fruit", "temp1","temp2","temp3"
## then test the columns created for numer of distinct values
## split the columns until you have columns of 
## subjects, properties, values, and metrics (where metrics
## are the units defined for the values)

## In this case, the subject is State/Strawberries -- 
## strawberries grown reported by state.

## When using separate_wider_delim() when you don't know the 
## number of columns the function will return,
## use the "too_many" and "too_few" parameters to set up 
## the function.  Generally, setting both parameters
## to "error" will produce helpful error messages.

  strwb_census <- strwb_census |>
  separate_wider_delim(  cols = `Data Item`,
                         delim = ",",
                         names = c("Fruit",
                                 "temp1",
                                 "temp2",
                                 "temp3"),
                         too_many = "error",
                         too_few = "align_start"
                       )

## Test the columns for the number of distinct values.
## for example:
##
# a <- strwb_census |> distinct(Fruit)
## The Fruit column only has one value: STRAWBERRIES the 
## subject under investigation.
##
## Remember - the value in single-value columns
## are often needed for Labels on tables and plots.
##
## Testing the temp1 column guides the next step.
# a <- strwb_census |> distinct(temp1)
## The "temp1" column has 4 distinct values
##
##    " ORGANIC - OPERATIONS WITH SALES"
##    " ORGANIC - PRODUCTION"           
##    " ORGANIC - SALES"                
##    " ORGANIC"  
##
##  (Note the leading space in each string -- 
##       which is fixed below.)
##
##  You can see that this column needs to be split between
##  "organic" and the properties "OPERATIONS WITH SALES", 
##  "PRODUCTION" and "SALES",  
##    using " - " as the column delimiter.
##
##  The column "prop_acct" contains the properties,
##   which are are accounting metrics related to
##   strawberry growing operations.


############################################
## split temp1 into crop_type, Prop_acct

strwb_census <- strwb_census |>
  separate_wider_delim(  cols = temp1,
                         delim = " - ",
                         names = c("crop_type",
                                 "prop_acct"),
                         too_many = "error",
                         too_few = "align_start"
                       )

## Once again, test the columns to plan your next step.
##
# a <- strwb_census |> distinct(crop_type)
## Column "crop_type' has single value  "organic"

# a <- strwb_census |> distinct(prop_acct)

## 
## The stringss in the "prop_acct" column are row labels
## for values reported in the "Values" column.  

##    "OPERATIONS WITH SALES"
##    "PRODUCTION"           
##    "SALES"               
##    "NA"   

## Note that the NA is in a row where the value 
## is labeled in another column.
##

############################################
## trim the strings
## you can see which columns contain string values that need
## to have leading or trailing spaces that need to be trimmed.


# glimpse(strwb_census)

strwb_census$crop_type <- str_trim(strwb_census$crop_type, side = "both")

strwb_census$temp2 <- str_trim(strwb_census$temp2, side = "both")

strwb_census$temp3 <- str_trim(strwb_census$temp3, side = "both")



#############################################
## split temp2 into market_type, measure

##
## The temp2 column requires a different logic.
## 

## start by looking at the unique entries in the temp2 column.

# a <- strwb_census |> distinct(temp2)
# 
# temp2
# 1  NA                                    
# 2 " MEASURED IN CWT"                     
# 3 " MEASURED IN $"                       
# 4 " FRESH MARKET - OPERATIONS WITH SALES"
# 5 " FRESH MARKET - SALES"                
# 6 " PROCESSING - OPERATIONS WITH SALES"  
# 7 " PROCESSING - SALES"   

## temp2 contains data for three separate columns
## 
##   All Strawberries  (is this a Total?)
##   Fresh Market
##   Processing
##
##  To understand these labels see 
##     "Strawberries: An Economic Assessment of the Feasibility
##      of Providing Multiple-Peril Crop Insurance",
##        prepared by Economic Research Service, USDA
##             for the Federal Crop Insurance Corporation
##                  October 31, 1994
## 

```

<!-- ## Create a "Fresh Market" column -->

```{r, include=FALSE}
#| label: create a fresh market column
#| echo: false
#| eval: true

## make a copy of the temp2 column named `Fresh Market`.
strwb_census <- strwb_census |> mutate(`Fresh Market` = temp2, .after = temp2)

## Remove cells in `Fresh Market` column 
##   that begin "MEASURED"
strwb_census$`Fresh Market` <- strwb_census$`Fresh Market` |> str_replace( "^MEA.*", "")

## Remove cells in `Fresh Market` column 
##   that begin "PROCESSING" 
strwb_census$`Fresh Market` <- strwb_census$`Fresh Market` |> str_replace( "^P.*", "")

## substitute a space for NA in `Fresh Market` column
strwb_census$`Fresh Market`[is.na(strwb_census$`Fresh Market`)] <- ""  

## in temp2 column, remove cells that begin "FRESH"
strwb_census$temp2 <- strwb_census$temp2 |> str_replace("^F.*", "")

## Now fix the entries in the `Fresh Market` column
##   Remove "FRESH MARKET - " from the cells
strwb_census$`Fresh Market` <- strwb_census$`Fresh Market` |> str_replace("^FRESH MARKET - ", "")


```

<!-- ## Create a "Process Market" column -->

```{r, include=FALSE}
#| label: make process market column
#| echo: false

## Make a copy of temp2 named `Process Market`
strwb_census <- strwb_census |> mutate(`Process Market` = temp2, .after = temp2)

## remove `Process Market` cells beginning "MEASURED"
strwb_census$`Process Market` <-  strwb_census$`Process Market` |> str_replace("^MEA.*", "")

## substitute space for NA in `Process Market` column
strwb_census$`Process Market`[is.na(strwb_census$`Process Market`)] <- ""

## In temp2, remove cells that begin "PROCESSING"
strwb_census$temp2 <- strwb_census$temp2 |> str_replace("^P.*", "")

## In `Processing Market`, remove "PROCESSING - " from cells
strwb_census$`Process Market` <-  strwb_census$`Process Market` |> str_replace("PROCESSING - ", "") 


```

<!-- Remove NA's from prop_acct, temp2, and temp3 -->

```{r, include=FALSE}
#| label: remove NAs
#| echo: false

## substitute a space for NA in prop_acct column
strwb_census$prop_acct[is.na(strwb_census$prop_acct)] <- "" 

## substitute a space for NA in temp2 column
strwb_census$temp2[is.na(strwb_census$temp2)] <- "" 

## substitute a space for NA in temp2 column
strwb_census$temp3[is.na(strwb_census$temp3)] <- "" 


```

<!-- Combine temp2 with temp3 to create Metric column -->

<!-- remove  -->

<!-- relocate columns -->

```{r, include=FALSE}
#| label: final cleanup
#| echo: false


strwb_census <- strwb_census |> unite(temp2, temp3, col="Metric", sep="")

## Now fix the entries in the Metric column
##   Remove "MEASURED IN " from the cells
strwb_census$Metric <- strwb_census$Metric |> str_replace("MEASURED IN ", "")

## move Metric to the end
strwb_census <- strwb_census |> relocate(Metric, .before = Domain)

strwb_census <- strwb_census |> relocate(`Process Market`, .before = Metric)

strwb_census <- strwb_census |> rename(Totals = prop_acct)

#drop_one_value_col(strwb_census)


```

<!-- ## The Value column transformation -->

```{r, include=FALSE}
#| label: define functions dcomma and footnote finder
#| echo: false
#| warning: false
#| message: false
#| eval: true

## remove commas from numbers
## fix footnotes

## basic tools

## start by getting the Values column so you can work on it 

vals <- strwb_census$Value

## note where vals goes in the environment.

## tools -- 2 choices  base R, and stringr package

## BaseR -- Piping??


g1 <- sub(",", "", vals)
# vals[1:20]
# g1[1:20]


g2 <- gsub(",", "", vals)
# vals[1:20]
# g2[1:20]


## stringr - str_replace(), str_replace_all()

## LOOK -- see ref for stingr pkg
a <- vals |> str_detect(",")

# vals[1:20]
# a[1:20]

## Still strings!!

b <- vals |> str_replace(",", "")
# vals[1:20]
# b[1:20]

c <- vals |> str_replace_all(",", "")
# vals[1:20]
# c[1:20]

## Now notice what happens when the
## the strings of digits are cast to numerics.

## for example
c <- as.numeric(c)
# c[1:20]


### remove commas from Value entries
dcomma <- function(c){
  x_new <- as.numeric(gsub(",", "", c))
  return(x_new)
}



#########################################  footnotes

## finds single uppor case Character in parens in s2
## e.g. "(D)"

## To fine the location and value of the footnotes

v <- strwb_census$Value


## find the footnote locations
## fn_i: locations 
fn_i <- v |> str_detect("^\\([:upper:]\\)$") ## returns


## dcomma returns numbers and NA's
v1 <- dcomma(v)

## locations of NA's
na_i <- is.na(v1)

## Demonstration that the locations of the footnotes
## are the same as the locations of the NA's

# length(v) == sum(na_i == fn_i)

## update dcomma()
## Integrate transformation of the values column and 
## reporting the footnote values.


dcomma <- function(c){
  suppressWarnings({
  xnew = as.numeric(gsub(",", "", c))
  fns = unique(c[is.na(xnew)])
  vtran = list("new_vec" = xnew, "footnotes" = fns)
  return(vtran)
  })
}

 
v_trns <- dcomma(v)
 

 a <- v_trns$new_vec
 # a[1:20]
 
 # v_trns$footnotes
 

```

<!-- NOTE -->

<!-- These plots were in early versions of the classroom notes. -->

<!-- In final versions of this document, they should be included in the EDA section. Note that the code will not run without editing. -->

<!-- ## plots   -->

```{r, include=FALSE}
#| label: plot 1
#| echo: false
#| eval: false

plot1_data <- strawberry |> 
  select(c(Year, State, Category, Value)) |> 
  filter((Year == 2021) & (Category == "ORGANIC - OPERATIONS WITH SALES"))

plot1_data$Value <- as.numeric(plot1_data$Value)

plot1_data <- plot1_data |> arrange(desc(Value))

ggplot(plot1_data, aes(x=reorder(State, -Value), y=Value)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(x = "States", y = "Count",
title ="Number of Organic Strawberry operations with Sales in 2021")


```

```{r, include=FALSE}
#| label: plot 2
#| echo: false
#| eval: false


##
plot2_data <- strawberry |> 
  select(c(Year, State, Category, Item, Value)) |> 
  filter((Year == 2021) & 
           (Category == "ORGANIC - SALES") & 
           (Item == "MEASURED IN $") & 
           (Value != "(D)"))


plot2_data$Value <- as.numeric(gsub(",", "", plot2_data$Value))

plot2_data <- plot1_data |> arrange(desc(Value))

ggplot(plot2_data, aes(x=reorder(State, -Value), y=Value)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(x = "States", y = "Sales",
title ="Organic Strawberry Sales ($) in 2021")


```

## Jiaxin's EDA for strawberry data

## Initial questions:

Initial questions about strawberries, the data, and about the work:
 
 (1) What is the degree of harmfulness of each chemical, which chemicals are more harmful and which are relatively harmless? 
 
 (2) How are strawberries sold in different states and in different years? What about the use of chemicals? 
 
 (3) Whether there is a correlation between different variables, e.g., whether strawberry sales are related to the year or the state in which they are located, etc. 
<!-- should be labeled and captioned. -->

<!-- ## chemicals -->

<!-- ### carcinogens from WHO list -->

<!-- #### updated -->

<!-- [list from WHO?](https://ggle.io/6Eys) -->

<!-- Arsenic: A Group 1 carcinogen, or conclusive cause of cancer   -->

<!-- Ethylene oxide: A Group 1 carcinogen   -->

<!-- Lindane: A Group 1 carcinogen   -->

<!-- 2,3,7,8-tetrachlorodibenzo-p-dioxin (TCDD): A Group 1 carcinogen   -->

<!-- Diazinon: Classified as "probably carcinogenic"   -->

<!-- Glyphosate: Classified as "probably carcinogenic"    -->

<!-- Malathion: Classified as "probably carcinogenic"   -->

<!-- #### -->

<!-- Now produce two data frames using strwb_survey. -->

<!-- One will have market data similar to strwb_census -->

<!-- The second will have data about chemicals used by  -->

<!-- strawberry growers. -->

<!-- Use the survey market data to enrich the data you already have -->

<!-- from the census.   -->

<!-- Use the chemical data to explore the use of carcinogens to grown strawberries.  Use the World Health Organization's -->

<!-- list of carcinogens.  These lists site specific chemicals.  -->

<!-- For exampe: these chemicals are on WHO's list of known carcinogens: captafol, ethylene dibromide, glyphosate, malathion, diazinon and dichlorophenyltrichloroethane (DDT).   Note that the WHO lists are not specific to strawberries.  You must use available data resources to determine which are used in strawberry growing. -->

<!-- ## Clean and organize strwb_survey -->

```{r, include=FALSE}
#| label: strwb_survey preliminary exploration
#| echo: false
#| eval: true

# glimpse(strwb_survey)

## find strwb_survey columns that should be split into columns
## 

## this section will produce tables listing
## the variables in the columns of strwb_survey
## remove the table your won't use

# distinct(strwb_survey[,3]) |> kable()
# c4 <- distinct(strwb_survey[,4])
# 
# c6 <- distinct(strwb_survey[,6])
# c7 <- distinct(strwb_survey[,7])
# c8 <- distinct(strwb_survey[,8])

# c3 |> kable()
# c4 |> kable()
# 
# c6 |> kable()
# c7 |> kable()
# c8 |> kable()


```

<!-- ## Column analyses -->

<!-- ### Period -->

```{r, include=FALSE}
#| label: period column 
#| echo: false

per_c <- strwb_survey |> select(Period) |> distinct()
per_c <- unlist(per_c)


## the Period column denotes
## three periods for data collection
##    marketing year
##    year
##    year - Aug Forecast
##



```

<!-- data item -->

```{r, include=FALSE}
#| label: data item analysis
#| echo: false


## columns need descriptive names

strwb_survey <- strwb_survey |> 
  separate_wider_delim(  cols = `Data Item`,
                         delim = ",",
                         names = c("temp1",
                                 "temp2",
                                 "temp3",
                                 "temp4"),
                         too_many = "error",
                         too_few = "align_start"
                       )

strwb_survey <- strwb_survey |>
  separate_wider_delim(  cols = temp1,
                         delim = " - ",
                         names = c("temp1a",
                                 "temp1b"),
                         too_many = "error",
                         too_few = "align_start"
                       )


# a <- strwb_survey |> distinct(temp1a)
## temp1a contains only 1 value -- STRAWBERRIES

a <- strwb_survey |> distinct(temp1a)

## It appears that strwb_survey intermingles
## market data rows and chemical data rows, with
## the indicators for the differet category of rows
## being in the "Domain" column.
## 

## let's examing the domain column -- begining lby
## separting with 
## splitting Domain into columns




```

<!-- Domain -->

```{r, include=FALSE}
#| label: domain
#| echo: false
#| eval: true



strwb_survey <- strwb_survey |>
  separate_wider_delim(  cols = Domain,
                         delim = ",",
                         names = c("temp22",
                                 "temp23"),
                         too_many = "error",
                         too_few = "align_start"
                       )


t22 <- unique(strwb_survey$temp22)

t23 <- unique(strwb_survey$temp23)

## Now separate the first column of the Domain Category.
## This iss allow the Chemical rows to be identified
## easity



```

<!-- Domain Category -->

```{r, include=FALSE}
#| label: Domain Category
#| echo: false
#| eval: true


strwb_survey <- strwb_survey |>
  separate_wider_delim(  cols = `Domain Category`,
                         delim = ",",
                         names = c("temp42",
                                 "temp43",
                                 "temp44",
                                 "temp45"),
                         too_many = "error",
                         too_few = "align_start"
                       )


## temp22 or temp42 or both == CHEMICAL 
##  else the row contains market data



strwb_survey_chem <- strwb_survey |> filter((temp22 == "CHEMICAL") | (temp42 == "CHEMICAL"))

strwb_survey_mkt <- strwb_survey |> filter(!((temp22 == "CHEMICAL") | (temp42 == "CHEMICAL")))


```

<!-- drop 1-value columns strawberry chems  -->

```{r, include=FALSE}
#| label: 1-value cols
#| echo: false
#| eval: true

chem1 <-  drop_one_value_col(strwb_survey_chem)

# chem1 |> kable(caption = "1-value columns dropped")

chem1 <- setdiff(colnames(strwb_survey_chem), chem1$col_name)

strwb_survey_chem <- strwb_survey_chem |> select(all_of(chem1))

```

<!-- mkt -->

```{r}
#| label: mkt 
#| echo: false
#| eval: true

mkt1 <- drop_one_value_col(strwb_survey_mkt)

# mkt1 |> kable(caption = "droping 1-value cols - mkt")

mkt1 <- setdiff(colnames(strwb_survey_mkt), mkt1$col_name)

strwb_survey_mkt <- strwb_survey_mkt |> select(all_of(mkt1))


```

<p style="page-break-before: always">

</p>


## References 

### Here are references that helps me while I'm writing. 
(Besides, thanks for Aidan's help about understanding variables, dealing with missing values, etc.)

[NASS help](https://quickstats.nass.usda.gov/tutorials)

[Quick Stats
Glossary](https://quickstats.nass.usda.gov/src/glossary.pdf)

[Quick Stats Column
Definitions](https://quickstats.nass.usda.gov/param_define)

[stats by
subject](https://www.nass.usda.gov/Statistics_by_Subject/index.php?sector=CROPS)

for EPA number lookup [epa
numbers](https://archive.epa.gov/pesticides/chemicalsearch/chemical/foia/web/html/128810.html)

[Active Pesticide Product Registration Informational
Listing](https://ordspub.epa.gov/ords/pesticides/f?p=APPRIL_PUBLIC:2::::::)

pc number input [pesticide chemical
search](https://ordspub.epa.gov/ords/pesticides/f?p=chemicalsearch:1)

[toxic chemical dashboard](https://comptox.epa.gov/dashboard/)

[ACToR -- Aggregated Computational Toxicology
Resource](https://cfpub.epa.gov/si/si_public_record_report.cfm?Lab=NCCT&dirEntryId=209598)

[comptox
dashboard](https://comptox.epa.gov/dashboard/chemical/details/DTXSID0020315)

[pubChem](https://pubchem.ncbi.nlm.nih.gov/)




## Investigating toxic pesticides


References that are helpful here:

[start here with chem PC
code](https://ordspub.epa.gov/ords/pesticides/f?p=chemicalsearch:1)

[step 2](https://ordspub.epa.gov/ords/pesticides/f?p=113:1::::RP,17,1::)
to get label (with warnings) for products using the chemical

[International Chemical safety
cards](https://www.ilo.org/dyn/icsc/showcard.home)

[Pesticide Product and Label
System](https://ordspub.epa.gov/ords/pesticides/f?p=113:1::::RP,17,1::)

[Search by
Chemical](https://ordspub.epa.gov/ords/pesticides/f?p=113:17::::::)

[CompTox Chemicals Dashboard](https://comptox.epa.gov/dashboard/)

[Active Pesticide Product Registration Informational
Listing](https://ordspub.epa.gov/ords/pesticides/f?p=APPRIL_PUBLIC:2::::::)

[OSHA chemical database](https://www.osha.gov/chemicaldata)

[Pesticide Ingredients](http://npic.orst.edu/ingred/)

[NPIC Product Research Online (NPRO)](http://npic.orst.edu/NPRO/)

[Databases for Chemical
Information](http://npic.orst.edu/ingred/cheminfo.html)

[Pesticide Active Ingredients](http://npic.orst.edu/ingred/active.html)

[TSCA Chemical Substance Inventory](https://www.epa.gov/tsca-inventory)

[glyphosate](https://ordspub.epa.gov/ords/pesticides/f?p=CHEMICALSEARCH:3::::1,3,31,7,12,25:P3_XCHEMICAL_ID:2478)


```{r, include=FALSE}
# View(strwb_census)
# View(strwb_survey)
# View(strwb_survey_mkt)
# View(strwb_survey_chem)
```

```{r, include=FALSE}
unique(strwb_survey_mkt$temp1b)
unique(strwb_survey_mkt$temp2)
unique(strwb_survey_mkt$temp3)
unique(strwb_survey_mkt$temp4)
unique(strwb_survey_mkt$temp42)

```

```{r, include=FALSE}

unique(strwb_survey_chem$temp4)

unique(strwb_survey_chem$temp23)

# unique(strwb_survey_chem$temp43)

```

## Matching processes

### First, I need to match different chemicals' PC codes, CAS numbers and their harm ranks. 

To be more specific,

(1) The numbers following the equals sign in these chemical substance
    records typically represent CAS numbers (Chemical Abstracts Service
    Registry Numbers). CAS numbers are a standardized numbering system
    used to uniquely identify chemical substances. Each CAS number is
    unique and serves the purpose of ensuring accurate identification
    and retrieval of information about a specific chemical substance.
    These numbers do not typically convey direct toxicity or hazard
    meanings; instead, they are used for the unique identification of
    chemicals in scientific and legal documents;

(2) In the list of 171 chemical substance records, there are 89 unique
    CAS numbers. The reason there are not 171 unique CAS numbers is that
    some chemicals share the same CAS number. This can happen when
    multiple substances have similar chemical structures or active
    ingredients and are therefore assigned the same CAS number. CAS
    numbers are meant to uniquely identify chemical substances, but
    similar or equivalent substances may share the same CAS number in
    some cases.

Then, In order to match the subsequent PC codes with their corresponding
CAS numbers, all the PC numbers in the "Data Items" column are first
extracted into a new separate column variable using regular expressions.


```{r}
library(dplyr)
library(writexl)

strwb_survey_chem_CAS <- strwb_survey_chem |>
  mutate(chemical_name = sub(".*\\((.*?)\\s=.*", "\\1", temp43),
         PCcodes = sub(".*=\\s(.*?)\\)", "\\1", temp43)) |>
  select(Year, State, temp23, chemical_name, PCcodes, Value) 

# colnames(strwb_survey_chem_CAS)
# View(strwb_survey_chem_CAS)
# write_xlsx(strwb_survey_chem_CAS, path = "strwb_survey_chem_CAS.xlsx")
```





## Data visualization about the usage of the four main groups of chemicals

Data visualization was used here to initially explore the use of the four main groups of chemicals("FUNGICIDE", "HERBICIDE", "INSECTICIDE","OTHER") in different states in different years:

(1) Here I created line plots using ggplot2 to visualize the proportion
    of different chemicals usage over time (from 2016 to 2018 to 2019
    to 2021) for four different states (CALIFORNIA, FLORIDA, OREGON,
    WASHINGTON). The proportion is defined as the count of a specific
    kind of chemical divided by the sum of the counts of the four types
    of chemicals: 'FUNGICIDE', 'HERBICIDE', 'INSECTICIDE' and 'OTHER';

(2) 'OREGON' and 'WASHINGTON' have data for 2016 but not for other
    years, it is likely due to a lack of data availability for these
    states in the subsequent years (2018, 2019, and 2021) in the data
    set. In other words, the data set we are working with might not
    include information for those states in those specific years,
    resulting in no data points for those combinations."

```{r}
unique(strwb_survey_chem_CAS$Year)
unique(strwb_survey_chem_CAS$State)
unique(strwb_survey_chem_CAS$temp23)
```


```{r}
library(ggplot2)
library(dplyr)

states_of_interest <- c("CALIFORNIA", "FLORIDA", "OREGON", "WASHINGTON")
filtered_data <- strwb_survey_chem_CAS %>%
  filter(State %in% states_of_interest)

proportion_data <- filtered_data %>%
  group_by(Year, State) %>%
  summarise(
    fungicide_count = sum(temp23 == " FUNGICIDE"),
    total_count = sum(temp23 %in% c(" FUNGICIDE", " HERBICIDE", " INSECTICIDE", " OTHER"))
  ) %>%
  mutate(proportion = round(fungicide_count / total_count, 3))

```


```{r}
library(ggplot2)
library(dplyr)

states_of_interest <- c("CALIFORNIA", "FLORIDA", "OREGON", "WASHINGTON")
filtered_data <- strwb_survey_chem_CAS %>%
  filter(State %in% states_of_interest)

proportion_data <- filtered_data %>%
  group_by(Year, State) %>%
  summarise(
    fungicide_count = sum(temp23 == " FUNGICIDE"),
    total_count = sum(temp23 %in% c(" FUNGICIDE", " HERBICIDE", " INSECTICIDE", " OTHER"))
  ) %>%
  mutate(proportion = round(fungicide_count / total_count, 3))

ggplot(proportion_data, aes(x = Year, y = proportion, color = State)) +
  geom_line() +
  geom_point(aes(label = sprintf("%.3f", proportion)), size = 3) +
  labs(
    x = "Year",
    y = "the proportion",
    title = "The Proportion of the Utility of Fungicide"
  ) +
  theme_minimal() +
  theme(legend.position = "top")

states_of_interest <- c("CALIFORNIA", "FLORIDA", "OREGON", "WASHINGTON")
filtered_data <- strwb_survey_chem_CAS %>%
  filter(State %in% states_of_interest)

proportion_data <- filtered_data %>%
  group_by(Year, State) %>%
  summarise(
    fungicide_count = sum(temp23 == " HERBICIDE"),
    total_count = sum(temp23 %in% c(" FUNGICIDE", " HERBICIDE", " INSECTICIDE", " OTHER"))
  ) %>%
  mutate(proportion = round(fungicide_count / total_count, 3))

ggplot(proportion_data, aes(x = Year, y = proportion, color = State)) +
  geom_line() +
  geom_point(aes(label = sprintf("%.3f", proportion)), size = 3) +
  labs(
    x = "Year",
    y = "the proportion",
    title = "The Proportion of the Utility of HERBICIDE"
  ) +
  theme_minimal() +
  theme(legend.position = "top")

states_of_interest <- c("CALIFORNIA", "FLORIDA", "OREGON", "WASHINGTON")
filtered_data <- strwb_survey_chem_CAS %>%
  filter(State %in% states_of_interest)

proportion_data <- filtered_data %>%
  group_by(Year, State) %>%
  summarise(
    fungicide_count = sum(temp23 == " INSECTICIDE"),
    total_count = sum(temp23 %in% c(" FUNGICIDE", " HERBICIDE", " INSECTICIDE", " OTHER"))
  ) %>%
  mutate(proportion = round(fungicide_count / total_count, 3))

ggplot(proportion_data, aes(x = Year, y = proportion, color = State)) +
  geom_line() +
  geom_point(aes(label = sprintf("%.3f", proportion)), size = 3) +
  labs(
    x = "Year",
    y = "the proportion",
    title = "The Proportion of the Utility of INSECTICIDE"
  ) +
  theme_minimal() +
  theme(legend.position = "top")

states_of_interest <- c("CALIFORNIA", "FLORIDA", "OREGON", "WASHINGTON")
filtered_data <- strwb_survey_chem_CAS %>%
  filter(State %in% states_of_interest)

proportion_data <- filtered_data %>%
  group_by(Year, State) %>%
  summarise(
    fungicide_count = sum(temp23 == " OTHER"),
    total_count = sum(temp23 %in% c(" FUNGICIDE", " HERBICIDE", " INSECTICIDE", " OTHER"))
  ) %>%
  mutate(proportion = round(fungicide_count / total_count, 3))

lineplot4 <- ggplot(proportion_data, aes(x = Year, y = proportion, color = State)) +
  geom_line() +
  geom_point(aes(label = sprintf("%.3f", proportion)), size = 3) +
  labs(
    x = "Year",
    y = "the proportion",
    title = "The Proportion of the Utility of OTHER"
  ) +
  theme_minimal() +
  theme(legend.position = "top")
lineplot4

```






Here, we could quickly check that there is data missing, definitely.
And take the 2018 year for an example:
```{r}
desired_year <- 2018
desired_state <- "OREGON" 

subset_data <- subset(strwb_survey_chem_CAS, Year == desired_year & State == desired_state)
print(subset_data)

desired_year <- 2018
desired_state <- "WASHINGTON" 

subset_data <- subset(strwb_survey_chem_CAS, Year == desired_year & State == desired_state)
# print(subset_data)

```

```{r}
strwb_survey_mkt <- strwb_survey_mkt[!grepl("\\(D\\)", strwb_survey_mkt$Value), ]
strwb_survey_mkt$Value <- as.numeric(strwb_survey_mkt$Value)

strwb_survey_mkt_grouped <- strwb_survey_mkt |>
  group_by(Year, State, temp1b) |>
  summarise(
    average_value = mean(Value)
  )

# View(strwb_survey_mkt_grouped)
```

## Exploring different chemicals' harm ranks

In the following, I will match the PC codes of the chemical substances
in the existing dataset to their corresponding CAS numbers, and then
match them to the hazard classes of the corresponding chemical
substances, with the help of the EXCEL datasheet provided by the
professor, which contains CAS numbers and hazard classes.

And about the harms of different chemicals with different CAS numbers,
we can learn from the file that prof shared that "INDEX. CLASS IFICATION
OF PESTICIDE AC TIVE INGREDIENTS: Ia = Extremely hazardous; Ib= Highly
hazardous; II=Moderately hazardous; III=Slightly hazardous; U = Unlikely
to present acute hazard in normal use; FM =Fumigant, not classified; O =
Obsolete as pesticide, not classified."

```{r}
cas_and_harm <- readxl::read_xlsx("CAS.xlsx")
# View(cas_and_harm)

```

```{r}
cas <- as.vector(unique(strwb_survey_chem_CAS$PCcodes))

PCcodes <- cas[nchar(cas) == 6 & grepl("^[0-9]+$", cas)]
PCcodes <- as.data.frame(PCcodes)
PCcodes$cas_valid <- c("23564-05-8", "15299-99-7", "38641-94-0", "70901-12-1", "13356-08-6", "153233-91-1", NA, "40487-42-1", "52315-07-8", "36734-19-7", "131929-60-7", NA, "42874-03-3", "70630-17-0", NA, "525-79-1", "57754-85-5", "67892-31-3", "112281-77-3", "99129-21-2", "161050-58-4", "108168-76-9", "60207-90-1", NA, "71751-41-2", "155569-91-8", "39148-24-8", "116714-46-6", "81777-89-1", "39515-41-8", "188425-85-6", "158062-67-0", "128639-02-1", "100784-20-1", "131860-33-8", "82657-04-3", "8000-78-0", "119446-68-3", "78587-05-0", "77182-82-2", "88671-89-0", "68694-11-1", "91465-08-6", "124-07-2", NA, "76674-21-0", "334-48-5", "81406-37-3", "66332-96-5", NA, "95737-68-1", "103361-09-7", "87674-68-8", "137497-61-1", NA, "122836-35-5", NA, "138261-41-3", "96489-71-3", "145701-23-1", "141517-21-7", "134098-61-6", "907204-31-3", "400882-07-7", "15708-41-5", NA, "146659-78-1", NA, "69327-76-0", "53112-28-0", "121552-61-2", "203313-25-1","180409-60-3") 
# View(PCcodes)



```

```{r}
library(dplyr)
strwb_survey_chem_CAS <- merge(strwb_survey_chem_CAS, PCcodes, by = "PCcodes")
strwb_survey_chem_CAS <- merge(strwb_survey_chem_CAS, cas_and_harm, by = "cas_valid")
# View(strwb_survey_chem_CAS)
# write_xlsx(strwb_survey_chem_CAS, path = "data_ with_cas_and_harms.xlsx")
```


## Dealing with the missing values

Handling of missing values in the "Value" variable: Calculate the mean
value of the "Value" variable from valid values other than "(D)" and
assign this calculated mean value to "(D)", "(NA)" and "(Z)". 
(Thanks for Aidan's inspiration here!)

```{r}
non_numeric_values <- c("(D)", "(NA)", "(Z)")
strwb_survey_chem_CAS$Value[strwb_survey_chem_CAS$Value %in% non_numeric_values] <- NA
mean_value <- mean(as.numeric(strwb_survey_chem_CAS$Value), na.rm = TRUE)
strwb_survey_chem_CAS$Value[is.na(strwb_survey_chem_CAS$Value)] <- mean_value
strwb_survey_chem_CAS$Value <- as.numeric(strwb_survey_chem_CAS$Value)
strwb_survey_chem_CAS$Value <-round(strwb_survey_chem_CAS$Value, digits = 3)

head(strwb_survey_chem_CAS)
#colnames(strwb_survey_chem_CAS)
#unique(strwb_survey_chem_CAS$harm_rank)
#unique(strwb_survey_chem_CAS$Year)
#unique(strwb_survey_chem_CAS$State)
```



```{r}
library(ggplot2)
ggplot(data = strwb_survey_chem_CAS, aes(x = Year, fill = harm_rank)) +
  geom_bar(position = "stack") +
  facet_wrap(~ State) +
  labs(title = "Chemical Harm Situations by State and Year",
       x = "Year",
       y = "Count of Chemical Harm Rank") +
  scale_fill_manual(values = c("Ib" = "red", "II" = "pink", "III" = "orange", "U" = "yellow")) +
  theme_classic()

```
From this, it can be seen that the harm ranks of chemicals used in different states in different years are predominantly II(Moderately hazardous) and U(Unlikely to present acute hazard in normal use.) different years is predominantly level 2 and level 5, which means that the hazards of the chemicals used between 2016 and 2021 are relatively acceptable.



## Data distributions and regression models

### exploring different states' situations about the harm ranks of chemicals

In addition, although the above analysis of the hazards of chemical substances has been carried out on a granular basis for each chemical substance, there are other factors that may have an impact on the harmlessness of different chemical substances in practice, such as the dosage used, whether strawberries are washed carefully, and so on.



### explore data in data frame "strwb_survey_chem" and "strwb_survey_mkt"

Here I want to explore data in data frame "strwb_survey_chem",
especially I'm curious about the relationships among variables like
"State", "Year", "Value". So I deal with the missing values in column
"Value" first, then I could try to create some visualization plots and
fit some models as below. This means that 

```{r}
non_numeric_values <- c("(D)", "(NA)", "(Z)")
strwb_survey_chem$Value[strwb_survey_chem$Value %in% non_numeric_values] <- NA
mean_value <- mean(as.numeric(strwb_survey_chem$Value), na.rm = TRUE)
strwb_survey_chem$Value[is.na(strwb_survey_chem$Value)] <- mean_value
strwb_survey_chem$Value <- as.numeric(strwb_survey_chem$Value)
strwb_survey_chem$Value <-round(strwb_survey_chem$Value, digits = 3)
# View(strwb_survey_chem)

hist(strwb_survey_chem$Value)

ggplot(data = strwb_survey_chem, aes(x=State, y=Value, color = State)) +
  geom_boxplot() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
    title = "Value situations of different states"
  )

ggplot(data = strwb_survey_chem, aes(x=Year, y=Value, color = State)) +
  geom_point() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
    title = "Value situations of different states in different years"
  )
```

Then I try to explore whether the factors like state, year, chemical
types could influence the Value, so I fit a generalized linear
model(model1) and linear model(model2) below.

But in both outputs, many coefficients has relatively large standard
deviation, and the t-stat indicates I fail to reject the Null
Hypothesis(All the coefficients are equal to zero), so these two model
don't perform well. In the future, I will try more model to explore
which model could fit well.

It turns out that the variable data are over dispersed so there are many
outliers and the usual linear models don't perform well.

```{r}
library(rstanarm)

model1 <- stan_glm(Value~Year + State + temp23, data=strwb_survey_chem, refresh=0)
summary(model1)
model2 <- lm(Value~Year + State + temp23, data=strwb_survey_chem)
summary(model2)

```



The last part of my EDA is about exploring data in the "strwb_survey_mkt", to 
explore the relationships among "state", "year", "Value", etc.
```{r}
mean_value <- mean(as.numeric(strwb_survey_mkt$Value), na.rm = TRUE)
strwb_survey_mkt$Value[is.na(strwb_survey_mkt$Value)] <- mean_value
strwb_survey_mkt$Value <-round(strwb_survey_mkt$Value, digits = 3)
# View(strwb_survey_mkt)


model3 <- stan_glm(Value~Year + State, data=strwb_survey_mkt, refresh=0)
summary(model1)

model4 <- lm(Value~Year + State, data=strwb_survey_mkt)
summary(model2)



hist(strwb_survey_mkt$Value)

ggplot(data = strwb_survey_mkt, aes(x=State, y=Value, color = State)) +
  geom_boxplot() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
    title = "Value situations of different states"
  )

ggplot(data = strwb_survey_mkt, aes(x=Year, y=Value, color = State)) +
  geom_point() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
    title = "Value situations of different states in different years"
  )

```
